---
title: "Zoe Thesis Analysis Code"
author: "Zoe Klotz"
date: "2026-03-22"
output: html_document
---

```{r setup, include=FALSE}
#load packages
library(dplyr)
library(ggplot2)
library(readr)
library(fixest)

getwd()
#insert your own working directory 
setwd([YOUR WORKING DIRECTORY])

#load data
Hospital_ER_DF <- read_csv("Hospital_Emergency_Department_Discharges_(SPARCS_De-Identified)__Potentially_Preventable_Emergency_Visit_(PPV)_Rates_by_Patient_Zip_Code__Beginning_2011_20260322.csv")

#view
glimpse(Hospital_ER_DF)
names(Hospital_ER_DF)

#Filter out non-NYC variables from Panel 
panel <- Hospital_ER_DF %>%
  filter(grepl("^[0-9]+$", `Patient Zipcode`)) %>%
  mutate(`Patient Zipcode` = sprintf("%05d", as.numeric(`Patient Zipcode`))) %>%
  filter(
    substr(`Patient Zipcode`, 1, 3) %in% c(
      "100", "101", "102", "103", "104",
      "111", "112", "113", "114", "116"
    )
  ) %>%
  filter(!is.na(`Risk Adjusted Rate Per 100 People`)) %>%
  rename(ppv_rate = `Risk Adjusted Rate Per 100 People`)

#see zip codes in panel (check to make sure it filtered correctly)
unique(panel$`Patient Zipcode`)

#set treated zipcode- UES (add 10044 to include Roosevelt Island)
treated_zips <- c("10021", "10028", "10128")

#for robustness checks later
treated_zips_ext <- c("10021", "10028", "10128", "10044")


```


```{r exploratory-comparison-treatment-groups-compared-to-all-Manhattan}

# Get mean pre-2017 PPV by Manhattan ZIP
baseline_candidates <- panel %>%
  filter(`Discharge Year` < 2017) %>%
  filter(substr(`Patient Zipcode`, 1, 3) == "100" | `Patient Zipcode` == "10128") %>%
  group_by(`Patient Zipcode`) %>%
  summarise(mean_ppv = mean(ppv_rate, na.rm = TRUE), .groups = "drop") %>%
  mutate(group = ifelse(`Patient Zipcode` %in% treated_zips, "Treated", "Control"))

baseline_candidates

# Visualize treated vs all other Manhattan ZIPs
ggplot(baseline_candidates, aes(x = group, y = mean_ppv)) +
  geom_boxplot() +
  geom_jitter(width = 0.1) +
  labs(
    title = "Baseline Mean PPV: Treated ZIPs vs All Manhattan ZIPs",
    x = "",
    y = "Mean PPV (Pre-2017)"
  )
```
```{r find-zipcodes-for-control}
# Compute baseline pre-treatment PPV for Manhattan candidate ZIPs
baseline <- panel %>%
  filter(`Discharge Year` < 2017) %>%
  filter(substr(`Patient Zipcode`, 1, 3) == "100" | `Patient Zipcode` == "10128") %>%
  group_by(`Patient Zipcode`) %>%
  summarise(mean_ppv = mean(ppv_rate, na.rm = TRUE), .groups = "drop")

# Visualize all Manhattan ZIPs with treated ZIPs marked in red
ggplot(baseline, aes(x = mean_ppv)) +
  geom_histogram(bins = 30, fill = "gray", color = "black") +
  geom_vline(
    data = baseline %>% filter(`Patient Zipcode` %in% treated_zips),
    aes(xintercept = mean_ppv),
    color = "red",
    linewidth = 1.2
  ) +
  labs(
    title = "Baseline PPV Distribution (Manhattan ZIPs)",
    x = "Mean PPV (Pre-2017)",
    y = "Count"
  )

# Visualize ZIP-level baseline PPV with labels
ggplot(baseline, aes(x = mean_ppv, y = reorder(`Patient Zipcode`, mean_ppv))) +
  geom_point() +
  geom_vline(
    data = baseline %>% filter(`Patient Zipcode` %in% treated_zips),
    aes(xintercept = mean_ppv),
    color = "red"
  ) +
  labs(
    title = "Baseline PPV by ZIP",
    x = "Mean PPV (Pre-2017)",
    y = "ZIP Code"
  )

# Based on these graphs, define matched controls using overlapping baseline PPV range
matched_controls <- baseline %>%
  filter(mean_ppv >= 11 & mean_ppv <= 14.5) %>%
  pull(`Patient Zipcode`) %>%
  setdiff(treated_zips)

# Print matched control ZIPs
matched_controls

# Table of treated and matched control ZIPs
comparison_tbl <- baseline %>%
  filter(`Patient Zipcode` %in% c(treated_zips, matched_controls)) %>%
  mutate(
    group = ifelse(`Patient Zipcode` %in% treated_zips, "Treated", "Control")
  ) %>%
  arrange(group, mean_ppv)

comparison_tbl

# Build final analysis dataframe
analysis_df <- panel %>%
  filter(`Patient Zipcode` %in% c(treated_zips, matched_controls)) %>%
  mutate(
    treated = ifelse(`Patient Zipcode` %in% treated_zips, 1, 0),
    group = ifelse(treated == 1, "Treated", "Control")
  )
```

comparison-between-treatment-and-control groups
```{r comparison-between-treatment-and-control groups}

comparison_tbl

# Box-and-whisker plot comparing treated and matched control ZIPs
ggplot(comparison_tbl, aes(x = group, y = mean_ppv)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.08, size = 2) +
  geom_text(aes(label = `Patient Zipcode`), hjust = -0.15, size = 3) +
  labs(
    title = "Baseline Mean PPV by Group",
    x = "",
    y = "Mean PPV (Pre-2017)"
  ) +
  coord_flip()


```
parallel-trend-analysis
```{r parallel-trend-analysis}

#visual analysis
trend <- analysis_df %>%
  filter(`Discharge Year` < 2017) %>%
  group_by(`Discharge Year`, group) %>%
  summarise(mean_ppv = mean(`ppv_rate`), .groups = "drop")

ggplot(trend, aes(x = `Discharge Year`, y = mean_ppv, color = group)) +
  geom_line(linewidth = 1.2) +
  geom_point() +
  labs(
    title = "Pre-Treatment Trends (Parallel trends check)"
  )
```
formal-pretrend-test
```{r formal-pretrend-test}

pre_model <- feols(
  ppv_rate ~ treated * factor(`Discharge Year`) | `Patient Zipcode`,
  data = analysis_df %>% filter(`Discharge Year` < 2017),
  cluster = ~`Patient Zipcode`
)

summary(pre_model)

#tidy, make exportable
library(broom)
trendtest <- tidy(pre_model)

#export to csv
write_csv(trendtest, "paralleltrendtest.csv")

#export table as html (optional)

#etable(
#  pre_model,
#  dict = c(
#    "factor(`Discharge Year`)2012" = "2012",
#    "factor(`Discharge Year`)2013" = "2013",
#    "factor(`Discharge Year`)2014" = "2014",
#    "factor(`Discharge Year`)2015" = "2015",
#    "factor(`Discharge Year`)2016" = "2016",
#    "treated:factor(`Discharge Year`)2012" = "Treated × 2012",
#    "treated:factor(`Discharge Year`)2013" = "Treated × 2013",
#    "treated:factor(`Discharge Year`)2014" = "Treated × 2014",
#    "treated:factor(`Discharge Year`)2015" = "Treated × 2015",
#    "treated:factor(`Discharge Year`)2016" = "Treated × 2016"
#  ),
#  file = "parallel_trend_test_pretty.html")

```

```{r preDiD-summarystats}
#Summary statistics
#overall average of both control and treatment groups 

overallavg <- analysis_df %>%
  summarise(
    mean_ppv = mean(ppv_rate, na.rm = TRUE),
    sd_ppv = sd(ppv_rate, na.rm = TRUE),
    min_ppv = min(ppv_rate, na.rm = TRUE),
    max_ppv = max(ppv_rate, na.rm = TRUE),
    n = n()
  )

overallavg

#control vs treated 

PreDiDsummary <- analysis_df %>%
  group_by(group) %>%
  summarise(
    mean_ppv = mean(ppv_rate, na.rm = TRUE),
    sd_ppv = sd(ppv_rate, na.rm = TRUE),
    min_ppv = min(ppv_rate, na.rm = TRUE),
    max_ppv = max(ppv_rate, na.rm = TRUE),
    n = n()
  )

PreDiDsummary

```

```{r DiD}
#Run DiD models
#prep /make sure it's clean  
analysis_df <- analysis_df %>%
  mutate(
    post = ifelse(`Discharge Year` >= 2017, 1, 0),
    ppv_rate = ppv_rate
  )

# 2-way fixed effect model 
did_model <- feols(
  ppv_rate ~ treated * post | `Patient Zipcode` + `Discharge Year`,
  data = analysis_df,
  cluster = ~`Patient Zipcode`
)

summary(did_model)

#run event model
event_model <- feols(
  ppv_rate ~ i(`Discharge Year`, treated, ref = 2016) |
    `Patient Zipcode` + `Discharge Year`,
  data = analysis_df,
  cluster = ~`Patient Zipcode`
)

summary(event_model)
iplot(event_model)

```

```{r sun-abraham}
#create treatment year if not there already
analysis_dfSuA <- analysis_df %>%
  mutate(
    treat_year = ifelse(treated == 1, 2017, 0)
  )

#run sun-abraham model
sa_model <- feols(
  ppv_rate ~ sunab(treat_year, `Discharge Year`) |
    `Patient Zipcode` + `Discharge Year`,
  data = analysis_dfSuA,
  cluster = ~`Patient Zipcode`
)

#results
summary(sa_model)

#plot
iplot(
  sa_model,
  main = "Sun-Abraham Event Study",
  xlab = "Event Time",
  ylab = "Effect on Risk-Adjusted PPV Rate"
)

```
```{r no2020plus}
#regualr DiD
#testing to try and see how Covid plays in 
analysis_pre_covid <- analysis_df %>%
  filter(`Discharge Year` <= 2019)


did_pre_covid <- feols(
  ppv_rate ~ treated * post | `Patient Zipcode` + `Discharge Year`,
  data = analysis_pre_covid,
  cluster = ~`Patient Zipcode`
)

summary(did_pre_covid)
#lowers effect, but does not disappear completely

#sunab event study no covid
#re-add treat_year
analysis_pre_covid <- analysis_pre_covid %>%
  mutate(
    treat_year = ifelse(treated == 1, 2017, 0)
  )

sa_pre_covid <- feols(
  ppv_rate ~ sunab(treat_year, `Discharge Year`) |
    `Patient Zipcode` + `Discharge Year`,
  data = analysis_pre_covid,
  cluster = ~`Patient Zipcode`
)


summary(sa_pre_covid)

iplot(
  sa_pre_covid,
  main = "No COVID: Sun-Abraham Event Study",
  xlab = "Event Time",
  ylab = "Effect on Risk-Adjusted PPV Rate"
)


```









alternative control method
```{r pretrend-match-slopeandmean}

#zip trends- calculate trend slope per zip (eg. the parallel trends graph)
zip_trends <- analysis_df %>%
  filter(`Discharge Year` < 2017) %>%
  group_by(`Patient Zipcode`) %>%
  summarise(
    slope = coef(lm(ppv_rate ~ `Discharge Year`))[2],
    mean_ppv = mean(ppv_rate),
    .groups = "drop"
  )

treated_profile <- zip_trends %>%
  filter(`Patient Zipcode` %in% treated_zips)

treated_slope <- mean(treated_profile$slope)
treated_mean  <- mean(treated_profile$mean_ppv)

#pull the zips closest to the treatment zips in both slope and mean ppv
matched_controls <- zip_trends %>%
  filter(!`Patient Zipcode` %in% treated_zips) %>%
  mutate(
    dist = sqrt(
      (mean_ppv - treated_mean)^2 +
      (slope - treated_slope)^2
    )
  ) %>%
  arrange(dist) %>%
  slice(1:5) %>%
  pull(`Patient Zipcode`)

matched_controls

#rebuild analysis dataframe only including these zips
analysis_df2 <- panel %>%
  filter(`Patient Zipcode` %in% c(treated_zips, matched_controls)) %>%
  mutate(
    treated = ifelse(`Patient Zipcode` %in% treated_zips, 1, 0),
    group = ifelse(treated == 1, "Treated", "Control")
  )

trend2 <- analysis_df2 %>%
  filter(`Discharge Year` < 2017) %>%
  group_by(`Discharge Year`, group) %>%
  summarise(mean_ppv = mean(`ppv_rate`), .groups = "drop")

ggplot(trend2, aes(x = `Discharge Year`, y = mean_ppv, color = group)) +
  geom_line(linewidth = 1.2) +
  geom_point() +
  labs(
    title = "Pre-Treatment Trends (Parallel trends check)"
  )


#somehow this graph is way worse with the algorithmically selected controls than the one's selected based on range

#table
pre_model2 <- feols(
  ppv_rate ~ treated * factor(`Discharge Year`) | `Patient Zipcode`,
  data = analysis_df2 %>% filter(`Discharge Year` < 2017),
  cluster = ~`Patient Zipcode`
)

summary(pre_model)


```


-----------------

STUDY 2- PPV across non-UES Q line (Effect of increased access to UES resources) 


```{r addcensusdata}
#Get ACS data
#install and load tidycensus package
library(tidycensus)
library(dplyr)
library(purrr)
library(tidyr)

# set your Census API key (only once ever)
## API key required to reproduce API calls.
#Go to https://www.census.gov/data/developers/data-sets/acs-5year.html to request a key 
census_api_key("[YOUR CENSUS API KEY HERE]", install = TRUE)

#Define variables
vars <- c(
  income = "B19013_001",          # median household income
  poverty_n = "B17001_002",       # number below poverty
  poverty_denom = "B17001_001",   # poverty denominator
  total_pop = "B03002_001",       #race table total pop
  white_nh = "B03002_003",
  black_nh = "B03002_004",
  asian_nh = "B03002_006",
  hispanic = "B03002_012",

#education (another measure of gentrification)
  male_ba       = "B15002_015",
  male_masters  = "B15002_016",
  male_prof     = "B15002_017",
  male_doctoral = "B15002_018",

  female_ba       = "B15002_032",
  female_masters  = "B15002_033",
  female_prof     = "B15002_034",
  female_doctoral = "B15002_035",
  edu_total = "B15002_001"
)

#Pull ACS 5-year estimates for years 2011 - 2022
years <- 2011:2022

#Get data from tables: BEWARE TAKES LONG TIME
acs_raw <- map_dfr(years, function(y) {
  get_acs(
    geography = "zcta",
    variables = vars,
    year = y,
    survey = "acs5"
  ) %>% 
    mutate(year = y)
})

#filter to NYC zip codes
acs_raw <- acs_raw %>%
  mutate(zcta = substr(GEOID, nchar(GEOID) - 4, nchar(GEOID))) %>%
  filter(substr(zcta, 1, 3) %in% c(
    "100","101","102","103","104","111","112","113","114","116"
  ))

#reshape table to wide
acs_wide <- acs_raw %>%
  select(GEOID, year, variable, estimate) %>%
  pivot_wider(
    names_from = variable,
    values_from = estimate
  )

#create clean ZCTA field
acs_wide <- acs_wide %>%
  mutate(zcta = substr(GEOID, nchar(GEOID) - 4, nchar(GEOID)))

#build ACS clean to add back to analysis: add college educated percentage and poverty rate
acs_clean <- acs_wide %>%
  mutate(
    poverty_rate = poverty_n / poverty_denom,
    college_educated = (male_ba + male_masters + male_prof + male_doctoral + female_ba + female_masters + female_prof + female_doctoral) / edu_total,             pct_white = white_nh / total_pop,
    pct_black = black_nh / total_pop,
    pct_hispanic = hispanic / total_pop) %>%
  select(
    zcta,
    year,
    income,
    poverty_rate,
    total_pop,
    college_educated, 
    pct_white, 
    pct_black, 
    pct_hispanic
  )

#view rows with NA data
acs_clean %>%
  filter(
    is.na(income) |
    is.na(poverty_rate) |
    is.na(total_pop) |
    is.na(college_educated)
  )

#and first 5 rows of full table
head(acs_clean)


```


Load in train lines + join with zctas, export train dataset w/ZCTAs
```{r extractZIPsfromTrains}

library(dplyr)
library(sf)
library(stringr)
library(tidyverse)

#Load MTA list (saved locally, downloaded from Open NYC data
subway_lines <- read_csv("~/Downloads/MTA_Subway_Stations_20251109.csv")

#isolate Q stops and set CRS
q_stops <- subway_lines %>%
  filter(str_detect(`Daytime Routes`, "Q")) %>%
  st_as_sf(coords = c("GTFS Longitude", "GTFS Latitude"), crs = 4326)

#load modZCTA
ZCTA_file <- read_csv("~/Desktop/MAthesis_data/Modified_Zip_Code_Tabulation_Areas_(MODZCTA)_20260329.csv")

#convert zcta to sf file
zcta_sf <- st_as_sf(ZCTA_file, wkt = "the_geom", crs = 4326)

#print(zcta_sf)
plot(st_geometry(zcta_sf))

#join trains + sf_ZCTA
q_with_zip <- st_join(q_stops, zcta_sf)


write_csv(q_with_zip, "Q_ZCTA.csv")

```




NAs in ACS dataset- dealing with missingness, looking at clustering 

```{r acsMissingness}

#identify where NAs exist in the dataset
acs_clean %>%
  summarise(across(everything(), ~mean(is.na(.))))


#How many per year? Are they concentrated in a single set of years or consistent across the timeframe? 
acs_clean %>%
  filter(
    is.na(income) |
    is.na(poverty_rate) |
    is.na(college_educated) |
    is.na(pct_black)
  ) %>%
  count(year)

#are they clustered geographically? 
acs_clean %>%
  filter(is.na(income) |
    is.na(poverty_rate) |
    is.na(college_educated) |
    is.na(pct_black)) %>%
  count(zcta)

#compare NA and non-NA rows
acs_clean %>%
  mutate(has_na = if_any(everything(), is.na)) %>%
  group_by(has_na) %>%
  summarise(
    mean_income = mean(income, na.rm = TRUE),
    mean_college = mean(college_educated, na.rm = TRUE),
    mean_black = mean(pct_black, na.rm = TRUE),
    n = n()
  )

#has_na rows always have mean income as a missing value; when NAs were removed from the dataset, there was nothing to average in the mean income column for the rows that had missing data. 

#which zctas is it? 
acs_clean %>%
  filter(is.na(income)) %>%
  distinct(zcta)

```



```{r MissingnessMap}

#create frame of missing incomes zctas and change zcta to character

acs_map <- acs_clean %>%
  mutate(missing_income = is.na(income)) %>% #create "missing income" binary column 
  mutate(zcta = as.character(zcta))

#join to modzcta, change zcta to character
zcta_map <- zcta_sf %>%
  mutate(MODZCTA = as.character(MODZCTA)) %>% 
  left_join(acs_map, by = c("MODZCTA" = "zcta"))


#check the data
#zcta_map$MODZCTA

sum(is.na(zcta_map$income))
nrow(zcta_map)

length(unique(zcta_sf$MODZCTA))
length(unique(acs_map$zcta))

intersect(
  unique(zcta_sf$MODZCTA),
  unique(acs_map$zcta)
) %>% length()


#extract ZIP codes that exist in the spatial dataset but do not exist in the ACS dataset
setdiff(unique(zcta_sf$MODZCTA), unique(acs_map$zcta))






```


```{r NAMap}

#create map of NAs
#collapse zcta_map to one row per zip
zcta_na_map <- zcta_map %>%
  group_by(MODZCTA) %>%
  summarise(
    missing_income = all(is.na(income)),
    geometry = first(the_geom)
  ) %>%
  ungroup()

#plot
library(ggplot2)

ggplot(zcta_na_map) +
  geom_sf(aes(fill = missing_income), color = "white", linewidth = 0.2) +
  scale_fill_manual(
    values = c("FALSE" = "lightblue", "TRUE" = "red"),
    name = "Missing Income"
  ) +
  labs(
    title = "ZIP Codes with Missing Income Data (ACS)",
    subtitle = "Red = no income data across all years"
  ) +
  theme_minimal()



```

```{r gentrification-def}
                                                                                                            
#define pre- and post-treatment years
pre_years <- 2011:2016
post_years <- 2017:2022

#compute pre- and post-treatment income/education/poverty averages for each ZCTA
acs_summary <- acs_clean %>%
  group_by(zcta) %>%
  summarise(
    income_pre = mean(income[year %in% pre_years], na.rm = TRUE),
    income_post = mean(income[year %in% post_years], na.rm = TRUE),

    college_pre = mean(college_educated[year %in% pre_years], na.rm = TRUE),
    college_post = mean(college_educated[year %in% post_years], na.rm = TRUE),

    poverty_pre = mean(poverty_rate[year %in% pre_years], na.rm = TRUE),
    poverty_post = mean(poverty_rate[year %in% post_years], na.rm = TRUE),

    .groups = "drop"
  )

#compute change over time
acs_summary <- acs_summary %>%
  mutate(
    d_income = income_post - income_pre,
    d_college = college_post - college_pre,
    d_poverty = poverty_post - poverty_pre
  )

#compute NYC average change
nyc_change <- acs_clean %>%
  summarise(
    income_pre = mean(income[year %in% pre_years], na.rm = TRUE),
    income_post = mean(income[year %in% post_years], na.rm = TRUE),

    college_pre = mean(college_educated[year %in% pre_years], na.rm = TRUE),
    college_post = mean(college_educated[year %in% post_years], na.rm = TRUE),

    poverty_pre = mean(poverty_rate[year %in% pre_years], na.rm = TRUE),
    poverty_post = mean(poverty_rate[year %in% post_years], na.rm = TRUE)
  ) %>%
  mutate(
    d_income_nyc = income_post - income_pre,
    d_college_nyc = college_post - college_pre,
    d_poverty_nyc = poverty_post - poverty_pre
  )

#compute change relative to NYC average
acs_summary <- acs_summary %>%
  mutate(
    rel_income = d_income - nyc_change$d_income_nyc,
    rel_college = d_college - nyc_change$d_college_nyc,
    rel_poverty = d_poverty - nyc_change$d_poverty_nyc
  )

#create relative gentrification index
acs_summary <- acs_summary %>%
  mutate(
    gentrification_index =
      scale(rel_income)[,1] +
      scale(rel_college)[,1] -
      scale(rel_poverty)[,1]
  )

        
acs_summary

#write.csv(acs_summary, "/Users/zo/Desktop/MAthesis_data/ACS_Gentscored_414.csv")


```

```{r define-intervention-area}

#note: NOT treatment zips. zips directly exposed to new infrastructure

# direct exposure ZIPs = ZIPs within 1 km of the new UES stations

ues_stops <- q_stops %>%
  filter(`Stop Name` %in% c("72 St", "86 St", "96 St"))

# project for distance/buffer work
ues_stops_proj <- st_transform(ues_stops, 6539)
zcta_sf_proj <- st_transform(zcta_sf, 6539)

# 1 km buffer around new stations
exp_q_buffer <- st_buffer(ues_stops_proj, dist = 1000) #1 km

# ZIPs intersecting the direct exposure zone
exp_q_zips_clean <- zcta_sf_proj[
  lengths(st_intersects(zcta_sf_proj, exp_q_buffer)) > 0,
]

direct_exposure_zips <- exp_q_zips_clean %>%
  st_drop_geometry() %>%
  pull(MODZCTA) %>%
  as.character() %>%
  unique() %>%
  sort()

#drop placeholdr zip 9999999
direct_exposure_zips <- direct_exposure_zips[direct_exposure_zips != "99999"]

direct_exposure_zips
length(direct_exposure_zips)


```


```{r}
# define full Q-corridor ZIP universe
# this is NOT yet the final treatment group

library(sf)
library(dplyr)

# project to a CRS with linear units for buffer calculations
q_stops_proj <- st_transform(q_stops, 2263)
zcta_sf_proj <- st_transform(zcta_sf, 2263)

# create 1 km buffer around all Q stops
q_corridor_buffer <- st_buffer(q_stops_proj, dist = 1500) #expanded to 1.5 km

# keep ZIPs that intersect the Q corridor buffer
q_corridor_zips <- zcta_sf_proj[
  lengths(st_intersects(zcta_sf_proj, q_corridor_buffer)) > 0,
]

# extract ZIP IDs
q_corridor_zip_ids <- q_corridor_zips %>%
  st_drop_geometry() %>%
  pull(MODZCTA) %>%
  as.character() %>%
  unique() %>%
  sort()

# drop placeholder ZIP
q_corridor_zip_ids <- q_corridor_zip_ids[q_corridor_zip_ids != "99999"]

# inspect
q_corridor_zip_ids
length(q_corridor_zip_ids)


```
ZIPS 

Manhattan (100xx & 10128)
10001 — Chelsea / Hudson Yards
10002 — Lower East Side
10003 — East Village / Union Square
10007 — Tribeca / Civic Center
10010 — Gramercy / Flatiron
10011 — Chelsea / Meatpacking District
10012 — SoHo / NoHo
10013 — Tribeca / Chinatown
10016 — Murray Hill / Kips Bay
10018 — Garment District / Midtown South
10019 — Midtown West
10021 — Upper East Side
10022 — Midtown East
10023 — Upper West Side
10028 — Upper East Side
10029 — East Harlem
10036 — Midtown / Times Square
10038 — Financial District
10065 — Upper East Side (Lenox Hill)
10075 — Upper East Side
10128 — Upper East Side (Carnegie Hill)
Brooklyn (112xx)
11201 — Brooklyn Heights / Downtown Brooklyn
11205 — Clinton Hill / Fort Greene
11210 — Flatbush / Midwood
11215 — Park Slope / Gowanus
11217 — Boerum Hill / Fort Greene
11218 — Kensington / Windsor Terrace
11223 — Gravesend
11224 — Coney Island
11225 — Crown Heights
11226 — Flatbush
11229 — Sheepshead Bay
11230 — Midwood
11235 — Brighton Beach
11238 — Prospect Heights / Clinton Hill


```{r matching-dfbuild}
#remove ues zips and determine treatment zips
#match treatment to control on gentrification scale pre and post-treatment years + match on income

#remove UES zips
treatment_zips <- q_corridor_zip_ids[
  !q_corridor_zip_ids %in% c("10021", "10028", "10029", "10065", "10128")
]

#view zips + num of zips
treatment_zips
length(treatment_zips)

# Build ZIP-level matching dataset from acs_summary, excluding UES zips, excluding NAs

match_df <- acs_summary %>%
  mutate(
    zcta = as.character(zcta),
    treated = ifelse(zcta %in% treatment_zips, 1, 0)
  ) %>%
  filter(
    !zcta %in% c("10021", "10028", "10029", "10065", "10128")
  ) %>%
  select(
    zcta,
    treated,
    income_pre,
    income_post,
    gentrification_index
  ) %>%
  filter(
    !is.na(income_pre),
    !is.na(income_post),
    !is.na(gentrification_index)
  )

match_df

#num of treated vs not
table(match_df$treated)

```


```{r matching-process}

library(MatchIt)

# Run matching
# start simple: nearest-neighbor on pre/post income + gentrification
m.out <- matchit(
  treated ~ income_pre + income_post + gentrification_index,
  data = match_df,
  method = "nearest",
  ratio = 1, 
  verbose = TRUE
)

summary(m.out)

#created matched dataset
matched_data <- match.data(m.out)

matched_data %>%
  count(treated) %>% 
  print()

```


```{r check-matching}

#compare match df to matched data; see how close control and treatment groups are after matchit

match_dfsum <- match_df %>%
  group_by(treated) %>%
  summarise(
    n = n(),
    mean_income_pre = mean(income_pre, na.rm = TRUE),
    mean_income_post = mean(income_post, na.rm = TRUE),
    mean_gentrification = mean(gentrification_index, na.rm = TRUE),
    .groups = "drop"
  )
#before
match_dfsum


matched_datasum <- matched_data %>%
  group_by(treated) %>%
  summarise(
    n = n(),
    mean_income_pre = mean(income_pre, na.rm = TRUE),
    mean_income_post = mean(income_post, na.rm = TRUE),
    mean_gentrification = mean(gentrification_index, na.rm = TRUE),
    .groups = "drop"
  )
#and after
matched_datasum



```






```{r extract-matched-zips}
# Extract matched ZIPs

matched_data <- match.data(m.out)

matched_treatment_zips <- matched_data %>%
  filter(treated == 1) %>%
  pull(zcta) %>%
  unique()

matched_control_zips <- matched_data %>%
  filter(treated == 0) %>%
  pull(zcta) %>%
  unique()

matched_treatment_zips
matched_control_zips

length(matched_treatment_zips)
length(matched_control_zips)



```
Build map of treatment and control zips
```{r zips_map}
#build map based on these zips (above)


library(dplyr)
library(ggplot2)
library(sf)

matched_treatment_zips <- as.character(matched_treatment_zips)
matched_control_zips   <- as.character(matched_control_zips)

#map it
map_df <- zcta_sf %>%
  mutate(MODZCTA = as.character(MODZCTA)) %>%
  mutate(
    group = case_when(
      MODZCTA %in% matched_treatment_zips ~ "Treatment",
      MODZCTA %in% matched_control_zips ~ "Control",
      TRUE ~ "Other"
    )
  )

ggplot(map_df) +
  geom_sf(aes(fill = group), color = "white", linewidth = 0.15) +
  labs(
    title = "Matched Treatment and Control ZIP Codes",
    subtitle = "Study 2 sample in NYC context",
    fill = "Group"
  ) +
  theme_minimal()



```

```{r build-panel}

#build panel

panel <- panel %>%
  mutate(`Patient Zipcode` = as.character(`Patient Zipcode`))

matched_treatment_zips <- as.character(matched_treatment_zips)
matched_control_zips <- as.character(matched_control_zips)

analysis_study2 <- panel %>%
  filter(`Patient Zipcode` %in% c(matched_treatment_zips, matched_control_zips)) %>%
  mutate(
    treated = ifelse(`Patient Zipcode` %in% matched_treatment_zips, 1, 0),
    group = ifelse(treated == 1, "Treated", "Control")
  )

```



```{r study2-ptrends}
#parallel trends

#plot trends
analysis_study2 %>%
  filter(`Discharge Year` < 2017) %>%
  group_by(`Discharge Year`, treated) %>%
  summarise(
    mean_ppv = mean(ppv_rate, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(`Discharge Year`, mean_ppv, color = factor(treated))) +
  geom_line() +
  geom_point()


```

```{r}

#formal parallel trends test
pre_model <- feols(
  ppv_rate ~ i(`Discharge Year`, treated, ref = 2016) | `Patient Zipcode`,
  data = analysis_study2 %>% filter(`Discharge Year` < 2017),
  cluster = ~`Patient Zipcode`
)


pre_model

```



Run event study for this
```{r}
#run event study for the overall dataset (no subsetting, cutting off years)

library(fixest)
#run event study model 

event_model_study2overall <- feols(
  ppv_rate ~ i(`Discharge Year`, treated, ref = 2016) |
    `Patient Zipcode` + `Discharge Year`,
  data = analysis_study2,
  cluster = ~`Patient Zipcode`
)

summary(event_model_study2overall)

#plot it 
iplot(
  event_model_study2overall,
  main = "Study 2: Event Study- Effect of Subway Expansion on PPV Rates (Overall)",
  xlab = "Event Time",
  ylab = "Effect (relative to 2016)",
  ref.line = 0
)

```






Next we examine heterogeneity within the sample, starting with income
```{r heterogeneity_ofZIPS_INCOME}

#look at income heterogeneity

library(dplyr)
#build diagnostic dataset from df.panel
diagnostic_df <- panel %>%
  mutate(`Patient Zipcode` = as.character(`Patient Zipcode`)) %>%
  left_join(
    acs_summary %>% mutate(zcta = as.character(zcta)),
    by = c("Patient Zipcode" = "zcta")
  )

#define treatment 
diagnostic_df <- diagnostic_df %>%
  mutate(
    treated = ifelse(`Patient Zipcode` %in% treatment_zips, 1, 0)
  )


#create bins (three bins)
diagnostic_df <- diagnostic_df %>%
  mutate(
    income_bin = ntile(income_pre, 3),
    income_bin = factor(income_bin, labels = c("Low", "Mid", "High"))
  )

#plot pretrends by income level
diagnostic_df %>%
  filter(`Discharge Year` < 2017) %>%
  group_by(`Discharge Year`, treated, income_bin) %>%
  summarise(mean_ppv = mean(ppv_rate), .groups = "drop") %>%
  ggplot(aes(`Discharge Year`, mean_ppv, color = factor(treated), group = treated)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2) +
  labs(title = "Income-Based Parallel Trends") +
  scale_color_manual(values = c("0" = "red", "1" = "blue")) +
  facet_wrap(~ income_bin) +
  labs(color = "Treated")+
  theme_minimal() +
  theme(panel.background = element_rect(fill = "white"))


#formal stratified 
pre_model_income <- feols(
  ppv_rate ~ i(`Discharge Year`, treated, ref = 2016) | `Patient Zipcode`,
  data = diagnostic_df %>% 
    filter(`Discharge Year` < 2017),
  cluster = ~`Patient Zipcode`,
  split = ~ income_bin
)
summary(pre_model_income)

```


closer look at the differences between the income brackets/ treated/not treated
```{r}
#calculate slop differences per bin
slope_diff <- diagnostic_df %>%
  filter(`Discharge Year` < 2017) %>%
  group_by(`Patient Zipcode`) %>%
  summarise(
    slope = coef(lm(ppv_rate ~ `Discharge Year`))[2],
    treated = first(treated),
    income_bin = first(income_bin),
    .groups = "drop"
  ) %>%
  group_by(income_bin, treated) %>%
  summarise(mean_slope = mean(slope), .groups = "drop") %>%
  tidyr::pivot_wider(names_from = treated, values_from = mean_slope) %>%
  mutate(
    `0` = as.numeric(`0`),
    `1` = as.numeric(`1`),
    diff = `1` - `0`
  )

#plot it
ggplot(slope_diff, aes(x = income_bin, y = diff, label = round(diff, 3))) +
  geom_col() +
  geom_text(vjust = -0.5) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(
    title = "Pre-Treatment Slope Differences by Income Group",
    x = "Income Group",
    y = "Slope Difference (Treated - Control)"
  ) +
  theme_minimal()



```

build map of low-med-high income zips (color by group)


```{r}
#build map
library(dplyr)
library(sf)
library(ggplot2)

# Define analysis sample ZIPs
analysis_zips <- unique(c(matched_treatment_zips, matched_control_zips))

# ZIP-level income + treatment
zip_income_groups <- diagnostic_df %>%
  select(`Patient Zipcode`, income_bin, treated) %>%
  distinct() %>%
  rename(MODZCTA = `Patient Zipcode`) %>%
  mutate(
    MODZCTA = as.character(MODZCTA),
    treated_label = ifelse(treated == 1, "Treatment", "Control")
  ) %>%
  filter(MODZCTA %in% analysis_zips)

# Full NYC map
nyc_map <- zcta_sf %>%
  mutate(MODZCTA = as.character(MODZCTA))

# Analysis ZIPs only
map_income <- nyc_map %>%
  filter(MODZCTA %in% analysis_zips) %>%
  left_join(zip_income_groups, by = "MODZCTA")

ggplot() +
  # Background NYC (all ZIPs)
  geom_sf(
    data = nyc_map,
    fill = "grey90",
    color = "white",
    linewidth = 0.1
  ) +
  # Control + Treatment ZIPs (colored by income)
  geom_sf(
    data = map_income,
    aes(fill = income_bin),
    color = "grey40",
    linewidth = 0.2
  ) +
  # Treatment ZIPs outlined
  geom_sf(
    data = map_income %>% filter(treated == 1),
    fill = NA,
    color = "black",
    linewidth = 0.9
  ) +
  scale_fill_manual(
    values = c(
      "Low" = "#1b9e77",
      "Mid" = "#d95f02",
      "High" = "#7570b3"
    )
  ) +
  labs(
    title = "Study 2 ZIP Codes by Income Group",
    subtitle = "Treatment ZIPs outlined; controls shown without outline",
    fill = "Income Group"
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    legend.position = "right"
  )

```
```{r eventmodel}

sa_model <- feols(
  ppv_rate ~ sunab(treat_year, `Discharge Year`) |
    `Patient Zipcode` + `Discharge Year`,
  data = analysis_dfSuA,
  cluster = ~`Patient Zipcode`
)

#results
summary(sa_model)

#plot
iplot(
  sa_model,
  main = "Sun-Abraham Event Study",
  xlab = "Event Time",
  ylab = "Effect on Risk-Adjusted PPV Rate"
)




```



Heterogeneity by distance


compute centroid distance to ues
```{r compute_dist_to_UES}

library(sf)
library(dplyr)

# standardize both to same CRS HERE 
ues_stops_proj <- st_transform(ues_stops, 2263)
zcta_sf_proj   <- st_transform(zcta_sf, 2263)

# ZIP centroids
zcta_centroids <- st_centroid(zcta_sf_proj)

# distance matrix
dist_matrix <- st_distance(zcta_centroids, ues_stops_proj)

# minimum distance
zcta_centroids$dist_to_q <- apply(dist_matrix, 1, min)

# keep ZIP ID + distance
zip_distance <- zcta_centroids %>%
  st_drop_geometry() %>%
  transmute(
    zcta = as.character(MODZCTA),
    dist_to_q = as.numeric(dist_to_q),
    dist_to_q_miles = as.numeric(dist_to_q) / 5280
  )

# merge onto diagnostic data
diagnostic_df <- diagnostic_df %>%
  mutate(`Patient Zipcode` = as.character(`Patient Zipcode`)) %>%
  left_join(
    zip_distance,
    by = c("Patient Zipcode" = "zcta")
  )

# create distance bins
diagnostic_df <- diagnostic_df %>%
  mutate(
    dist_bin = ntile(dist_to_q_miles, 3),
    dist_bin = factor(dist_bin, labels = c("Near", "Mid", "Far"))
  )

summary(diagnostic_df$dist_to_q_miles)
table(diagnostic_df$dist_bin, useNA = "ifany")

```

build map of binned distances (color = group)
```{r map-dist}

library(sf)
library(dplyr)
library(ggplot2)

# Define analysis sample ZIPs
analysis_zips <- unique(c(matched_treatment_zips, matched_control_zips))

# ZIP-level distance + treatment lookup
zip_distance_groups <- diagnostic_df %>%
  select(`Patient Zipcode`, dist_bin, treated) %>%
  distinct() %>%
  rename(MODZCTA = `Patient Zipcode`) %>%
  mutate(MODZCTA = as.character(MODZCTA)) %>%
  filter(MODZCTA %in% analysis_zips)

# Full NYC map
nyc_map <- zcta_sf %>%
  mutate(MODZCTA = as.character(MODZCTA))

# Analysis ZIPs only
map_distance <- nyc_map %>%
  filter(MODZCTA %in% analysis_zips) %>%
  left_join(zip_distance_groups, by = "MODZCTA")

# Plot: only treatment/control colored, treatment outlined
ggplot() +
  # Background NYC
  geom_sf(
    data = nyc_map,
    fill = "grey90",
    color = "white",
    linewidth = 0.1
  ) +
  # Analysis ZIPs colored by distance bin
  geom_sf(
    data = map_distance,
    aes(fill = dist_bin),
    color = "grey40",
    linewidth = 0.2
  ) +
  # Treatment ZIPs outlined
  geom_sf(
    data = map_distance %>% filter(treated == 1),
    fill = NA,
    color = "black",
    linewidth = 0.9
  ) +
  scale_fill_manual(
    values = c(
      "Near" = "#1b9e77",
      "Mid" = "#d95f02",
      "Far" = "#7570b3"
    )
  ) +
  labs(
    title = "Study 2 ZIP Codes by Distance Group",
    subtitle = "Treatment ZIPs outlined; controls shown without outline",
    fill = "Distance Group"
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    legend.position = "right"
  )





```


```{r dist_hetero_plot}
#plot
diagnostic_df %>%
  filter(`Discharge Year` < 2017) %>%
  group_by(`Discharge Year`, treated, dist_bin) %>%
  summarise(mean_ppv = mean(ppv_rate), .groups = "drop") %>%
  ggplot(aes(`Discharge Year`, mean_ppv, color = factor(treated), group = treated)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2) +
  labs(title = "Distance-Based Parallel Trends") +
  scale_color_manual(values = c("0" = "red", "1" = "blue")) +
  facet_wrap(~ dist_bin) +
  labs(color = "Treated")+
  theme_minimal() +
  theme(panel.background = element_rect(fill = "white"))



```
Looks like for near and mid zips, the trajectory becomes more parallel after 2014, while for far zips, they diverge pretty significantly at 2016. 



```{r distformaltest}

#formal test for dist
pre_model_dist <- feols(
  ppv_rate ~ i(`Discharge Year`, treated, ref = 2016) | `Patient Zipcode`,
  data = diagnostic_df %>% filter(`Discharge Year` < 2017),
  cluster = ~`Patient Zipcode`,
  split = ~ dist_bin
)

summary(pre_model_dist)

```

```{r dist_slopes}

#compute slopes per zip
slope_dist <- diagnostic_df %>%
  filter(`Discharge Year` < 2017) %>%
  group_by(`Patient Zipcode`) %>%
  summarise(
    slope = coef(lm(ppv_rate ~ `Discharge Year`))[2],
    treated = first(treated),
    dist_bin = first(dist_bin),
    .groups = "drop"
  )


#avg slopes by group
slope_summary_dist <- slope_dist %>%
  group_by(dist_bin, treated) %>%
  summarise(mean_slope = mean(slope), .groups = "drop") %>%
  pivot_wider(names_from = treated, values_from = mean_slope) %>%
  mutate(
    `0` = as.numeric(`0`),
    `1` = as.numeric(`1`),
    diff = `1` - `0`
  )

slope_summary_dist

#plot slope differences
ggplot(slope_summary_dist, aes(x = dist_bin, y = diff, label = round(diff, 3))) +
  geom_col() +
  geom_text(vjust = -0.5) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(
    title = "Pre-Treatment Slope Differences by Distance",
    x = "Distance Group",
    y = "Slope Difference (Treated - Control)"
  ) +
  theme_minimal()


```

Since there were clear signs of trends for certain groups becoming more parallel post 2013, and signs of near zips not being parallel at all, we're going to restrict to comparable geographic areas and stable pre-periods (eg. post 2013, old far and mid zips distance-wise)- testing increased access

```{r}

library(dplyr)
library(ggplot2)

# restrict sample: drop Near + keep later pre-period
analysis_df_final <- diagnostic_df %>%
  filter(
    dist_bin != "Near",
    `Discharge Year` >= 2013,
    `Discharge Year` < 2017   # keep pre-treatment only
  )

# create pre-trend plot
analysis_df_final %>%
  group_by(`Discharge Year`, treated, dist_bin) %>%
  summarise(mean_ppv = mean(ppv_rate, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(`Discharge Year`, mean_ppv, color = factor(treated))) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  facet_wrap(~ dist_bin) +
  labs(
    title = "Pre-Treatment Trends (Restricted Sample: Mid & Far, 2013–2016)",
    x = "Discharge Time",
    y = "Mean PPV Rate",
    color = "Treated"
  ) +
  theme_minimal()


```

Use far only as controls? EG: compare treated ZIPs to geographically distant ZIPs that exhibit similar pre-treatment trends
```{r}

analysis_df_far <- diagnostic_df %>%
  filter(
    dist_bin == "Far" | treated == 1,   # keep all treated + far controls
    `Discharge Year` >= 2013
  )%>%
  mutate(
    group = ifelse(treated == 1, "Treated", "Control"),
    post = ifelse(`Discharge Year` >= 2017, 1, 0),
    treat_year = ifelse(treated == 1, 2017, 0)
  )

# Quick check
analysis_df_far %>%
  count(treated, dist_bin)


#Pre-trend plot

analysis_df_far %>%
  filter(`Discharge Year` < 2017) %>%
  group_by(`Discharge Year`, treated) %>%
  summarise(
    mean_ppv = mean(ppv_rate, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(`Discharge Year`, mean_ppv, color = factor(treated))) +
  geom_line(linewidth = 1.1) +
  geom_point(size = 2) +
  labs(
    title = "Pre-Treatment Trends: Treated ZIPs vs Far Controls (2013–2016)",
    x = "Discharge Time",
    y = "Mean PPV Rate",
    color = "Treated"
  ) +
  theme_minimal()

#formal pretrend test
pre_model_far <- feols(
  ppv_rate ~ i(`Discharge Year`, treated, ref = 2016) | `Patient Zipcode`,
  data = analysis_df_far %>% filter(`Discharge Year` < 2017),
  cluster = ~`Patient Zipcode`
)

summary(pre_model_far)


```

Run DiD model 
```{r}

did_model_far <- feols(
  ppv_rate ~ treated * post | `Patient Zipcode` + `Discharge Year`,
  data = analysis_df_far,
  cluster = ~`Patient Zipcode`
)

summary(did_model_far)


```
DEEPLY insignificant results lmao 

Event study DiD
```{r}
library(fixest)
#run event study model 
event_model_far <- feols(
  ppv_rate ~ i(`Discharge Year`, treated, ref = 2016) |
    `Patient Zipcode` + `Discharge Year`,
  data = analysis_df_far,
  cluster = ~`Patient Zipcode`
)

summary(event_model_far)

#plot it 
iplot(
  event_model_far,
  main = "Event Study: Effect of Subway Expansion on PPV Rates",
  xlab = "Event Time",
  ylab = "Effect (relative to 2016)",
  ref.line = 0
)


#much smaller confidence intervals in pre-period vs post
# might be an issue with the data, choice of control 
```


```{r study2-SUNAB}
#sunab for far zips
sa_study2 <- feols(
  ppv_rate ~ sunab(treat_year, `Discharge Year`) |
    `Patient Zipcode` + `Discharge Year`,
  data = analysis_df_far,
  cluster = ~`Patient Zipcode`
)


summary(sa_study2)

iplot(
  sa_study2,
  main = "Sun-Ab Event Study (Study 2): Effect of Subway Expansion on PPV Rates (far zip codes)",
  xlab = "Year",
  ylab = "Effect (relative to 2016)",
  ref.line = 0
)

#significant for first year after intervention, then drops down to year 0 levels (not sustained effect)
#transitory effect

```






Restrict sample in different ways (run with near + 2013-2022)
```{r}
#NEAR ZIPS (Exclude 2011-2012)
#pretrend + DiD 
library(ggplot2)
library(dplyr)
library(fixest)

# Build analysis sample
analysis_near <- diagnostic_df %>%
  filter(
    `Discharge Year` >= 2013,
    `Discharge Year` <= 2022,
    `Patient Zipcode` %in% c(matched_treatment_zips, matched_control_zips)
  ) %>%
  mutate(
    post = ifelse(`Discharge Year` >= 2017, 1, 0), 
    treat_year = ifelse(treated == 1, 2017, 0)
  )

# Quick sanity checks
analysis_near %>% count(treated)
analysis_near %>% count(`Discharge Year`)

# Run DiD
did_near <- feols(
  ppv_rate ~ treated * post | `Patient Zipcode` + `Discharge Year`,
  data = analysis_near,
  cluster = ~`Patient Zipcode`
)

summary(did_near)

analysis_near %>%
  filter(`Discharge Year` < 2017) %>%
  group_by(`Discharge Year`, treated) %>%
  summarise(mean_ppv = mean(ppv_rate, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(`Discharge Year`, mean_ppv, color = factor(treated))) +
  geom_line() +
  geom_point() +
  labs(
    title = "Pre-Treatment Trends (Near + All Bins, 2013–2016)",
    color = "Treated"
  ) +
  theme_minimal()


#run sunab model on near zips

#event study
nearzips_eventmodel <- feols(
  ppv_rate ~ sunab(treat_year, `Discharge Year`) |
    `Patient Zipcode` + `Discharge Year`,
  data = analysis_near,
  cluster = ~`Patient Zipcode`
)

#results
summary(nearzips_eventmodel)

#plot
iplot(
  nearzips_eventmodel,
  main = "Near ZIPs: Sun-Ab Event Study",
  xlab = "Event Time",
  ylab = "Effect on Risk-Adjusted PPV Rate"
)






```


run by income 
-  just low income
-  just high income 
```{r}
#high income 

library(dplyr)
library(fixest)

# Build sample
analysis_high <- diagnostic_df %>%
  filter(
    `Discharge Year` >= 2013,
    `Discharge Year` <= 2022,
    `Patient Zipcode` %in% c(matched_treatment_zips, matched_control_zips),
    income_bin == "High"
  ) %>%
  mutate(
    post = ifelse(`Discharge Year` >= 2017, 1, 0),
    treat_year = ifelse(treated == 1, 2017, 0)
  )

# Sanity checks
analysis_high %>% count(treated)
analysis_high %>% count(`Discharge Year`)

# Run DiD
did_high <- feols(
  ppv_rate ~ treated:post | `Patient Zipcode` + `Discharge Year`,
  data = analysis_high,
  cluster = ~`Patient Zipcode`)

summary(did_high)

#pretrend
analysis_high %>%
  filter(`Discharge Year` < 2017) %>%
  group_by(`Discharge Year`, treated) %>%
  summarise(mean_ppv = mean(ppv_rate, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(`Discharge Year`, mean_ppv, color = factor(treated))) +
  geom_line() +
  geom_point() +
  labs(
    title = "Pre-Treatment Trends (High-Income ZIPs, 2013–2016)",
    color = "Treated"
  ) +
  theme_minimal()


#sunab event study
highinc_eventmodel <- feols(
  ppv_rate ~ sunab(treat_year, `Discharge Year`) |
    `Patient Zipcode` + `Discharge Year`,
  data = analysis_high,
  cluster = ~`Patient Zipcode`
)

#results
summary(highinc_eventmodel)

#plot
iplot(
  highinc_eventmodel,
  main = "High Income ZIPs: Sun-Ab Event Study",
  xlab = "Event Time",
  ylab = "Effect on Risk-Adjusted PPV Rate"
)


```
```{r}
#low income
library(dplyr)
library(fixest)

# Build sample
analysis_low <- diagnostic_df %>%
  filter(
    `Discharge Year` >= 2013,
    `Discharge Year` <= 2022,
    `Patient Zipcode` %in% c(matched_treatment_zips, matched_control_zips),
    income_bin == "Low"
  ) %>%
  mutate(
    post = ifelse(`Discharge Year` >= 2017, 1, 0),
    treat_year = ifelse(treated == 1, 2017, 0)

  )

# Sanity checks
analysis_low %>% count(treated)
analysis_low %>% count(`Discharge Year`)

# Run DiD
did_low <- feols(
  ppv_rate ~ treated * post | `Patient Zipcode` + `Discharge Year`,
  data = analysis_low,
  cluster = ~`Patient Zipcode`
)

summary(did_low)

#pretrend
analysis_low %>%
  filter(`Discharge Year` < 2017) %>%
  group_by(`Discharge Year`, treated) %>%
  summarise(mean_ppv = mean(ppv_rate, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(`Discharge Year`, mean_ppv, color = factor(treated))) +
  geom_line() +
  geom_point() +
  labs(
    title = "Pre-Treatment Trends (Low-Income ZIPs, 2013–2016)",
    color = "Treated"
  ) +
  theme_minimal()

#low inc event study
low_inc_eventmodel <- feols(
  ppv_rate ~ sunab(treat_year, `Discharge Year`) |
    `Patient Zipcode` + `Discharge Year`,
  data = analysis_low,
  cluster = ~`Patient Zipcode`
)

#results
summary(low_inc_eventmodel)

#plot
iplot(
  low_inc_eventmodel,
  main = "Low Income ZIPs: Sun-Ab Event Study",
  xlab = "Event Time",
  ylab = "Effect on Risk-Adjusted PPV Rate"
)


```

